Enhanced low-rank matrix estimation for simultaneous denoising and reconstruction of five-dimensional seismic data

### 2021-08-09 elrme5d/paper.tex
Mon Aug  9 16:40:12 CDT 2021

We focus first on the parameters that the proposed proximity operator requires. The parameter $\varphi_i$ \wen{should be}\dlo{is} a positive integer $(\varphi_i>0)$. \wen{To see its influence \old{in}\new{on} the proposed algorithm, we set ${n}=3$, ${q}=3$, ${\beta_0}=0.971$ and ${\beta_1}=0.977$, $\delta_0 = 3.75$ and $\delta_1 = 3.25$, ${b_0}=25$ and ${b_1}=23$\old{ by trial}, and we vary the value of $\varphi_i$ from $1$ to $5$ when estimating the low-rank signal matrix of the first synthetic data example. Figure \ref{fig:WindowsSizeφi} shows the SNRs diagram with different values of ${\varphi_i}$. We find that all values of ${\varphi_i}$ selected above $1$ weaken the \new{performance of the }proposed operator}\dlo{This parameter}\old{$\varphi_i$ produces some signal damages of data when it is greater than $1$}. To achieve the best quality of the result, we propose to set \old{$\varphi_0$ and $\varphi_1$ equal to $1$}\new{${\varphi_0}={\varphi_1}=1$} for any noisy and incomplete data to be estimated. \wen{In the same perspective, we keep the values of parameters mentioned including ${\varphi_0}={\varphi_1}=1$, and we show the strength of both ${\beta_0}$ and ${\beta_1}$ in Figure \ref{fig:RationalTransfertFunctionCoefficientbeta0and1}. From $0.900$ to $0.970$, ${\beta_0}$ helps to produce better results compared with ${\beta_1}$. In contrast, ${\beta_1}$ becomes better than ${\beta_0}$ from $0.980$ to $1$. Both parameters yield the same values of SNR around ${\beta_i = 0.974}$. But this value is not the best trade-off to achieve the performance of the proposed approach. Therefore, we need to adjust ${\beta_0}$ and ${\beta_1}$ around $0.970$ and $0.980$, respectively. After some trials we adopted ${\beta_0 = 0.971}$ and ${\beta_1 = 0.977}$ to estimate the signal matrix.} Usually, when $\beta_i$ is large, the smoothing level degrades the significant feature of the useful signal. If $\beta_i$ is too small, it generates extra noise in the data. Therefore, we usually choose $\beta_i \in \left[0, 1\right]$. As a general rule, $\beta_0$ should always be less than or equal to $\beta_1$ ($\beta_0 \leq \beta_1$) whatever the data to be restored is. \wen{Also}, as the iteration numbers and the noise levels increase, we should set $\beta_i$ by increasing gradually the values towards $1$. However, when the percentage of missing traces becomes high (from $10\% $ to $90\%$), we should select $\beta_i$ progressively from $1$ to $0$. The couples of parameters ($\varphi_0$, $\beta_0$) and ($\varphi_1$, $\beta_1$) are used to filter out the unwanted values that the singular values contain and the elements of the matrix to be estimated, respectively. \wen{This guideline is also verified for data \old{including}\new{with} curved events and/or complex structures like almost all field seismic data}. Then, we set the regularization parameters $\lambda_i$ as we proposed in the theory section. We use the regulating parameters called cooling factor $\delta_i$ to adjust $\lambda_i$. When the ratio of missing traces increases, the values of $\delta_i$ should decrease gradually. In contrast, when the noise level increases (from $0.1$ to $1$), the values of $\delta_i$ should increase gradually. The cooling factors $\delta_i$ provide us some flexibility to control the estimation of the low-rank matrix. The proposed regularization parameters $\lambda_i$ allow\old{s} exploiting sparsity along all dimensions of multidimensional seismic data. \wen{Figure \ref{fig:CoolingfactorsSigma0and1} shows the contribution of $\delta_0$ and $\delta_1$ in obtaining the result of the proposed strategy. As shown, from $0$ to $4$, the parameter $\delta_0$ (magenta line) aids at producing almost the same SNRs value. Its SNRs diagram seems to be invariable compared to that of $\delta_1$ (green line), which becomes stable from $20$ to $25$. Here,} to achieve a best trade-off with other parameters, we propose to set $\delta_0 \geq \delta_1$. \wen{Therefore, we have selected $\delta_0 = 3.75$ and $\delta_1 = 3.25$ when estimating the low-rank signal matrix having linear events. Relatively small values of $\delta_i$ can be selected for data, including curved events or complex structures.} Besides, the penalty convexity parameters $b_i$ should be greater than zero (${b_i} > 0 $). \wen{For this parameter, we analyze the simultaneous denoising and reconstruction quality of the ELRME algorithm by keeping all other parameters, while varying ${b_0}$ and ${b_1}$ from $1, 2, 3,..,$ to $25$. From Figure \ref{fig:PenaltyConvexityParametersb0b1}, the best values of both parameters can be selected around $23$. But we suggest setting ${b_0}\geq {b_1}$ for obtaining the best trad-off with the other parameters. Here, we selected ${b_0}=25$ and ${b_1}=23$ to recover the signal matrix. This analysis reveals that larger values of ${b_i}$ can be adopted for the restoration of data, including linear events. In contrast, in case of data including curved events, smaller values are demanded}. Note that, if ${b_i}$ is too small (${b_i} < 1e^{-10}$), there is no advantage in applying the arctangent penalty function, and the result is almost the same as applying the soft-thresholding function. It is important to know that when $\delta_i$ and ${b_i}$ are very large, it may degrade the amplitude of the estimated signal. ${b_0}$ and ${b_1}$ affect the sparsity of the singular values, and the elements of the noisy matrix, respectively. The rank ${n}$ needs to be set according to the number of different dip components, \old{with}\new{for} data \old{having}\new{with} linear events. For data containing curved events, ${n}$ needs to be relatively higher than the number of different dip components \citep{oropeza2010randomized}. \wen{Figure \ref{fig:rankn} highlights the effects of ${n}$ in our results for different input values. We adopted ${n}=3$ because of the best quality of the results. The 5D synthetic data used in this paper includes $3$ linear events. Thus, ${n} = 3$ confirms the linear events assumption. We display the influence of the damping factor for different values in Figure \ref{fig:dampingFactorq}. From this figure, we find that the damping operator of the proposed approach loses its ability when the parameter ${q}$ is very large. \old{The damping operator used in the DRR algorithm and that of the new algorithm behave, therefore, in the same way.} Based on this remark, ${q}$ should be chosen in the interval of ${q} \in \left[2, 3\right]$ for 5-D synthetic data and field data sets having linear or curved events as suggested in \cite{chen2016open} and \cite{chen2016simultaneous}}.

@@

\inputdir{./}
\multiplot{6}{rankn,dampingFactorq,WindowsSizeφi,RationalTransfertFunctionCoefficientbeta0and1,CoolingfactorsSigma0and1,PenaltyConvexityParametersb0b1}{width=0.3\textwidth}{Weight of each parameter. (a) rank constraint (n). (b) Damping factor (q). (c) Windows size ($\varphi_i$). (d) Rational transfert function coefficient ($\beta_i$). (e) Cooling factor ($\delta_i$). (f) Penalty convexity parameters ${b_i}$.}

>>>>

We focus first on the parameters that the proposed proximity operator requires. The parameter $\varphi_i$ \wen{should be}\dlo{is} a positive integer $(\varphi_i>0)$. \wen{To see its influence \old{in}\new{on} the proposed algorithm, we set ${n}=3$, ${q}=3$, ${\beta_0}=0.971$ and ${\beta_1}=0.977$, $\delta_0 = 3.75$ and $\delta_1 = 3.25$, ${b_0}=25$ and ${b_1}=23$\old{ by trial}, and we vary the value of $\varphi_i$ from $1$ to $5$ when estimating the low-rank signal matrix of the first synthetic data example. Figure \ref{fig:WindowsSizephi} shows the SNRs diagram with different values of ${\varphi_i}$. We find that all values of ${\varphi_i}$ selected above $1$ weaken the \new{performance of the }proposed operator}\dlo{This parameter}\old{$\varphi_i$ produces some signal damages of data when it is greater than $1$}. To achieve the best quality of the result, we propose to set \old{$\varphi_0$ and $\varphi_1$ equal to $1$}\new{${\varphi_0}={\varphi_1}=1$} for any noisy and incomplete data to be estimated. \wen{In the same perspective, we keep the values of parameters mentioned including ${\varphi_0}={\varphi_1}=1$, and we show the strength of both ${\beta_0}$ and ${\beta_1}$ in Figure \ref{fig:RationalTransfertFunctionCoefficientbeta0and1}. From $0.900$ to $0.970$, ${\beta_0}$ helps to produce better results compared with ${\beta_1}$. In contrast, ${\beta_1}$ becomes better than ${\beta_0}$ from $0.980$ to $1$. Both parameters yield the same values of SNR around ${\beta_i = 0.974}$. But this value is not the best trade-off to achieve the performance of the proposed approach. Therefore, we need to adjust ${\beta_0}$ and ${\beta_1}$ around $0.970$ and $0.980$, respectively. After some trials we adopted ${\beta_0 = 0.971}$ and ${\beta_1 = 0.977}$ to estimate the signal matrix.} Usually, when $\beta_i$ is large, the smoothing level degrades the significant feature of the useful signal. If $\beta_i$ is too small, it generates extra noise in the data. Therefore, we usually choose $\beta_i \in \left[0, 1\right]$. As a general rule, $\beta_0$ should always be less than or equal to $\beta_1$ ($\beta_0 \leq \beta_1$) whatever the data to be restored is. \wen{Also}, as the iteration numbers and the noise levels increase, we should set $\beta_i$ by increasing gradually the values towards $1$. However, when the percentage of missing traces becomes high (from $10\% $ to $90\%$), we should select $\beta_i$ progressively from $1$ to $0$. The couples of parameters ($\varphi_0$, $\beta_0$) and ($\varphi_1$, $\beta_1$) are used to filter out the unwanted values that the singular values contain and the elements of the matrix to be estimated, respectively. \wen{This guideline is also verified for data \old{including}\new{with} curved events and/or complex structures like almost all field seismic data}. Then, we set the regularization parameters $\lambda_i$ as we proposed in the theory section. We use the regulating parameters called cooling factor $\delta_i$ to adjust $\lambda_i$. When the ratio of missing traces increases, the values of $\delta_i$ should decrease gradually. In contrast, when the noise level increases (from $0.1$ to $1$), the values of $\delta_i$ should increase gradually. The cooling factors $\delta_i$ provide us some flexibility to control the estimation of the low-rank matrix. The proposed regularization parameters $\lambda_i$ allow\old{s} exploiting sparsity along all dimensions of multidimensional seismic data. \wen{Figure \ref{fig:CoolingfactorsSigma0and1} shows the contribution of $\delta_0$ and $\delta_1$ in obtaining the result of the proposed strategy. As shown, from $0$ to $4$, the parameter $\delta_0$ (magenta line) aids at producing almost the same SNRs value. Its SNRs diagram seems to be invariable compared to that of $\delta_1$ (green line), which becomes stable from $20$ to $25$. Here,} to achieve a best trade-off with other parameters, we propose to set $\delta_0 \geq \delta_1$. \wen{Therefore, we have selected $\delta_0 = 3.75$ and $\delta_1 = 3.25$ when estimating the low-rank signal matrix having linear events. Relatively small values of $\delta_i$ can be selected for data, including curved events or complex structures.} Besides, the penalty convexity parameters $b_i$ should be greater than zero (${b_i} > 0 $). \wen{For this parameter, we analyze the simultaneous denoising and reconstruction quality of the ELRME algorithm by keeping all other parameters, while varying ${b_0}$ and ${b_1}$ from $1, 2, 3,..,$ to $25$. From Figure \ref{fig:PenaltyConvexityParametersb0b1}, the best values of both parameters can be selected around $23$. But we suggest setting ${b_0}\geq {b_1}$ for obtaining the best trad-off with the other parameters. Here, we selected ${b_0}=25$ and ${b_1}=23$ to recover the signal matrix. This analysis reveals that larger values of ${b_i}$ can be adopted for the restoration of data, including linear events. In contrast, in case of data including curved events, smaller values are demanded}. Note that, if ${b_i}$ is too small (${b_i} < 1e^{-10}$), there is no advantage in applying the arctangent penalty function, and the result is almost the same as applying the soft-thresholding function. It is important to know that when $\delta_i$ and ${b_i}$ are very large, it may degrade the amplitude of the estimated signal. ${b_0}$ and ${b_1}$ affect the sparsity of the singular values, and the elements of the noisy matrix, respectively. The rank ${n}$ needs to be set according to the number of different dip components, \old{with}\new{for} data \old{having}\new{with} linear events. For data containing curved events, ${n}$ needs to be relatively higher than the number of different dip components \citep{oropeza2010randomized}. \wen{Figure \ref{fig:rankn} highlights the effects of ${n}$ in our results for different input values. We adopted ${n}=3$ because of the best quality of the results. The 5D synthetic data used in this paper includes $3$ linear events. Thus, ${n} = 3$ confirms the linear events assumption. We display the influence of the damping factor for different values in Figure \ref{fig:dampingFactorq}. From this figure, we find that the damping operator of the proposed approach loses its ability when the parameter ${q}$ is very large. \old{The damping operator used in the DRR algorithm and that of the new algorithm behave, therefore, in the same way.} Based on this remark, ${q}$ should be chosen in the interval of ${q} \in \left[2, 3\right]$ for 5-D synthetic data and field data sets having linear or curved events as suggested in \cite{chen2016open} and \cite{chen2016simultaneous}}.

@@

\inputdir{./}
\multiplot{6}{rankn,dampingFactorq,WindowsSizephi,RationalTransfertFunctionCoefficientbeta0and1,CoolingfactorsSigma0and1,PenaltyConvexityParametersb0b1}{width=0.3\textwidth}{Weight of each parameter. (a) rank constraint (n). (b) Damping factor (q). (c) Windows size ($\varphi_i$). (d) Rational transfert function coefficient ($\beta_i$). (e) Cooling factor ($\delta_i$). (f) Penalty convexity parameters ${b_i}$.}
